---
title: "ikarus cancer cell classification in R"
author: "Vedran Franke"
email: "vedran.franke@mdc-berlin.de"
date: "today"
output: 
  html_notebook:
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preliminaries

## Setup the conda environment

Here we install the appopriate python version, along with all
of the required R and python packages.

```{python, eval=FALSE}
conda create  -y -n ikar3.8 -python=3.8
conda install -y libgcc
conda install -y r-essentials r-base
conda install -y -c bioconda bioconductor-biocinstaller bioconductor-zellkonverter bioconductor-singlecellexperiment
conda install -y r-seurat
pip install git+https://github.com/BIMSBbioinfo/ikarus.git
```


# Load the conda environment in R

Start the R interpreter, load **reticulate**, and
load the conda environment.

**conda_path** should contain the path to the installed conda environment.

```{r load_reticulate}
library(reticulate)
conda_path = c("~/bin/Software/miniconda3/envs/ikar3.8")
use_condaenv(conda_path)
```


# Data preparation

## Download the ikarus model

```{r download_model}
trained_model_path = "https://github.com/BIMSBbioinfo/ikarus/raw/master/tutorials/out/core_model.joblib"

download.file(trained_model_path, "core_model.joblib")
```

### Download gene signatures

```{r download_signatures}
signatures_path = "https://github.com/BIMSBbioinfo/ikarus/raw/master/tutorials/out/signatures.gmt"
download.file(signatures_path, "signatures.gmt")
```

# Run Ikarus

## Download the adata

```{r download_data}
adata_path = "https://bimsbstatic.mdc-berlin.de/akalin/Ikarus/part_1/data/tirosh17_headneck/adata.h5ad"
download.file(adata_path, "tirosh_adata.h5ad")
```

## Read the h5ad into an anndata object

```{r read_anndata}
anndata = import("anndata")
adata = anndata$read_h5ad("tirosh_adata.h5ad")
```

## Load the trained ikarus model

```{r load_ikarus}
ikarus = import("ikarus")
model  = ikarus$classifier$Ikarus(
     
  signatures_gmt = file.path("signatures.gmt"), 
  out_dir="ikarus_path"
)
model$load_core_model("core_model.joblib")
```

## Predict the cell class

```{r predict_cell_class}
class_predict = model$predict(adata, "test")
```

## Look at the prediction results

```{r compare_results}
results = data.frame(
  true = adata$obs$tier_0,
  pred = class_predict
)
with(results, table(true, pred))

```


# MISC helpers

## Seurat to SingleCellExperiment

```{r seurat_to_singleCellExperiment, eval=FALSE}
sce = SingleCellExperiment(
  assays = list(counts = GetAssayData(seu, "counts", "RNA"),
  colData = DataFrame(seu@meta.data)
) 
```


## Convert **SingleCellExperiment** to **AnnData**


```{r, singlecellexperiment_to_anndata, eval=FALSE}
library(basilisk)
adata = basiliskRun(fun = function(sce) {
     # Convert SCE to AnnData:
     SCE2AnnData(sce)
     
}, env = conda_path, sce = sce)
```

## Process count data using ikarus

If the input **adata** object contains counts, the matrix
can be processed using ikarus **preprocess_adata** function.

If the data is already normalized, no preprocessing is required

**IMPORTANT** the data should not be scaled!

```{r process_data, eval=FALSE}
adata = ikarus$data$preprocess_adata(adata)
```



